---
title: "SenthilDhanapal_IS621_HW1"
author: "Senthil Dhanapal"
date: "September 24, 2016"
output: pdf_document
---
```{r, echo=FALSE,warning=FALSE}
#install.packages("corrplot")
library(corrplot)
require(corrplot)

```


```{r, echo=FALSE}
winsorize <- function(x, multiple=2.2)
{
  q <- quantile(x)
  iqr <- IQR(x)
  iqrAdjusted <- iqr*multiple
  
  rangeLow <- q['25%']-iqrAdjusted
  rangeHigh <- q['75%']+iqrAdjusted
  
  x[x<rangeLow] <- min(x[x>rangeLow])
  x[x>rangeHigh] <- max(x[x<rangeHigh])
  
  return(x)  
  
}


```

```{r, echo=FALSE}
#setwd("f:/temp")
setwd("C:/Senthil/MSDataAnalytics/Semester4/621/Week1")
dfMB <- read.csv("moneyball.csv",stringsAsFactors = FALSE)


```

#################################################################################################
#Variables Considered
#-------------------------------------------------------------
#TEAM_BATTING_H = Base Hits by batters (1B,2B,3B,HR)
#TEAM_BATTING_3B = Triples by batters (3B)
#TEAM_PITCHING_H = Hits allowed
#-------------------------------------------------------------
#################################################################################################
```{r, echo=FALSE}
dfMB <- dfMB[,c("Number.of.wins","Base.Hits.by.batters..1B.2B.3B.HR.", 
                "Triples.by.batters..3B.","Hits.allowed")]
#dfMB <- na.omit(dfMB)
colnames(dfMB) <- c('TARGET_WINS','TEAM_BATTING_H','TEAM_BATTING_3B','TEAM_PITCHING_H')


```

#################################################################################################
## Data Exploration using summary and plots
#-------------------------------------------------------------
#################################################################################################
```{r, echo=FALSE}
summary(dfMB)

###check distribution of data 
par(mfrow=c(2,2))

hist(dfMB$TARGET_WINS)
hist(dfMB$TEAM_BATTING_H)
hist(dfMB$TEAM_BATTING_3B)
hist(dfMB$TEAM_PITCHING_H)

boxplot(dfMB$TARGET_WINS, main="Number of wins")
boxplot(dfMB$TEAM_BATTING_H, main="Base Hits by batters (1B,2B,3B,HR)")
boxplot(dfMB$TEAM_BATTING_3B, main="Triples by batters (3B)")
boxplot(dfMB$TEAM_PITCHING_H, main="Hits allowed ")


```

#################################################################################################
# Handling Outliers using Winsorizing method
#-------------------------------------------------------------
#################################################################################################

```{r, echo=FALSE}

dfMB$TEAM_BATTING_H <- winsorize(dfMB$TEAM_BATTING_H)
dfMB$TEAM_BATTING_3B <- winsorize(dfMB$TEAM_BATTING_3B)
dfMB$TEAM_PITCHING_H <- winsorize(dfMB$TEAM_PITCHING_H)


```

#################################################################################################
## Data Exploration after winsorizing using plots
#-------------------------------------------------------------
#################################################################################################
```{r, echo=FALSE}

par(mfrow=c(2,2))

hist(dfMB$TARGET_WINS)
hist(dfMB$TEAM_BATTING_H)
hist(dfMB$TEAM_BATTING_3B)
hist(dfMB$TEAM_PITCHING_H)

boxplot(dfMB$TARGET_WINS, main="Number of wins")
boxplot(dfMB$TEAM_BATTING_H, main="Base Hits by batters (1B,2B,3B,HR)")
boxplot(dfMB$TEAM_BATTING_3B, main="Triples by batters (3B)")
boxplot(dfMB$TEAM_PITCHING_H, main="Hits allowed ")


```


#################################################################################################
#Check for relevancy and multicollinearity using Correllation between all variables
#-------------------------------------------------------------
#################################################################################################

```{r, echo=FALSE}
par(mfrow=c(1,1))
cor(dfMB, use="all.obs", method="pearson") 
M <- cor(dfMB)
corrplot(M, method="circle")


```

## Correllation betwen TEAM_BATTING_H and TEAM_BATTING_3B may lead to multicollinearity
## TEAM_BATTING_H has the best correllation with Target wins than any other predictors.
## TEAM_PITCHING_H has no correllation 
## However, keeping all 3 variables to build models
#-------------------------------------------------------------

#################################################################################################
# Build Models
#-------------------------------------------------------------
#################################################################################################

#Model1
```{r, echo=FALSE}
model1 <- lm(TARGET_WINS~TEAM_BATTING_H+TEAM_PITCHING_H+TEAM_BATTING_3B,dfMB)
summary(model1)
par(mfrow=c(2,2))
plot(model1)


```


#Model2
```{r, echo=FALSE}
model2 <- lm(TARGET_WINS~TEAM_BATTING_H+TEAM_PITCHING_H,dfMB)
summary(model2)
par(mfrow=c(2,2))
plot(model2)


```


#Model3
```{r, echo=FALSE}
model3 <- lm(TARGET_WINS^2~TEAM_BATTING_H+TEAM_PITCHING_H,dfMB)
summary(model3)
par(mfrow=c(2,2))
plot(model3)


```

#Model4
```{r, echo=FALSE}
model4 <- lm(TARGET_WINS~TEAM_BATTING_H,dfMB)
summary(model4)
par(mfrow=c(2,2))
plot(model4)


```



#################################################################################################
# Choosing Models
#-------------------------------------------------------------
#################################################################################################

```{r, echo=FALSE}

model <- c(1,2,3,4)
adjRSqrd <- c(0.1711,0.1689,0.1653,0.1461)
FStat <- c(157.5,232.2,226.3,390.3)
PValForFStat <- c('Significant','Significant','Significant','Significant')
ResidualVsFittedConstantVariation <- c('Not good,', 'Not good', 'Not good', 'Not good')
ResidualVsFittedCuve <- c('shows curve,', 'shows curve', 'shows curve', 'shows curve')
ResidualVsFittedCuveHeteroscedasticity  <- c('yes,', 'yes', 'yes', 'yes')
print(data.frame(model,adjRSqrd,FStat,PValForFStat,ResidualVsFittedConstantVariation,ResidualVsFittedCuve,ResidualVsFittedCuveHeteroscedasticity))


```

