---
title: "Homework 1"
author: "Group 1"
date: ''
output:
  pdf_document:
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
---

```{r, include = FALSE}
library(Hmisc)
library(xlsx)
library(knitr)
library(magrittr)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(e1071)
library(corrplot)
library(knitcitations)
library(bibtex)
```

```{r, include=FALSE, cache=TRUE}
download.file("https://github.com/ChristopheHunt/DATA-621-Group-1/blob/master/Data%20Dictionary.xlsx?raw=true", 
              destfile = "Data Dictionary HW1 - Group 1.xlsx", mode='wb')
data_dictionary <- read.xlsx("Data Dictionary HW1 - Group 1.xlsx", sheetIndex = 1)
file.remove(dir(getwd(), pattern = "Data Dictionary HW1 - Group 1", full.names = TRUE))
```


# Introduction

The ability to analyize and predict performance of a team using many dimensions is critical to competitive success in the professional Baseball industry. Therefore, we analyized the records of numerous professional baseball team from the years 1871 to 2006. Our hope is that the following report and the resulting predictive models will better inform the organization and assist in making data driven decisions moving forward. 

"The goal of a baseball team is to win more games than any other team. Since one team has very little control over the number of games other teams win, the goal is essentially to win as many games as possible. Therefore, it is of interest to measure the player's contribution to the team's wins." Grabiner, B. D. [^1] 

[^1]: (Grabiner, B. D. (n.d.). [The Sabermetric Manifesto.](http://seanlahman.com/baseball-archive/sabermetrics/sabermetric-manifesto/) Retrieved September 10, 2016 from http://seanlahman.com/baseball-archive/sabermetrics/sabermetric-manifesto/) 

# Statement of the Problem

The purpose of this report is to determine the batting, baserun, pitching, and fielding effects on a baseball team's ability to win.  

# Data Exploration  

Please note that each record has the performance of the team for the given year, with all of the statistics adjusted to match the performance of a 162 game season. 

The following table provides the descriptive statistics regarding our the available data set. You will find that 

Describe the size and the variables in the moneyball training data set. Consider that too much detail will cause a
manager to lose interest while too little detail will make the manager consider that you aren't doing your job. Some
suggestions are given below. Please do NOT treat this as a check list of things to do to complete the assignment.
You should have your own thoughts on what to tell the boss. These are just ideas.

```{r, echo=FALSE, results='asis'}
df <- read.csv("https://raw.githubusercontent.com/ChristopheHunt/DATA-621-Group-1/master/moneyball-training-data.csv")
colnames(df) <- mapvalues(as.vector(colnames(df)), 
                            from = str_trim(data_dictionary$VARIABLE.NAME..), to = as.vector(data_dictionary$DEFINITION))
digits <- c(1) 

descriptive <- describe(df %>% select(-`Identification Variable (do not use)`), 
                        descript = "Descriptive Statistics", digits = digits) 

for (i in seq(1:length(descriptive))){for (j in c(6,9,12)){names(descriptive[[i]]$counts)[j] <- paste0(names(descriptive[[i]]$counts)[j]," freq")
    }
     descriptive[[i]]$counts <- (descriptive[[i]]$counts[-c(4,7,8:10,11)])
     
     descriptive[[i]]$counts[7] <- round(sapply(df[i+1], function(x) median(x, na.rm=TRUE)), digits = digits)
     names(descriptive[[i]]$counts)[7] <- "median"
     
     descriptive[[i]]$counts[8] <- round(sapply(df[i+1], function(x) sd(x, na.rm=TRUE)), digits = digits)
     names(descriptive[[i]]$counts)[8] <- "sd"
     
     descriptive[[i]]$counts[9] <- round(sapply(df[i+1], function(x) skewness(x, na.rm=TRUE)), digits = digits)
     names(descriptive[[i]]$counts)[9] <- "skewness"
     
     descriptive[[i]]$counts[10] <- round(sapply(df[i+1], function(x) kurtosis(x, na.rm=TRUE)), digits = digits)
     names(descriptive[[i]]$counts)[10] <- "kurtosis"
     
     descriptive[[i]]$counts <- (descriptive[[i]]$counts[c(1:4,7:8,5:6,9:10)])
     
}

latex(descriptive, file = '')
```

\newpage

c. Is the data correlated to the target variable (or to other variables?)

Correlation Matrix

```{r, results='asis', echo = FALSE }

#or <- rcorr(as.matrix(df %>% select(-`Identification Variable (do not use)`)), type="pearson")
x <- corrplot(cor(df %>% select(-`Identification Variable (do not use)`), use="complete.obs"),
          method="square",  tl.col = "black", tl.cex=.5, type="lower"
       )
  
```

\newpage

# Data Preparation  

Describe how you have transformed the data by changing the original variables or creating new variables. If you
did transform the data or create new variables, discuss why you did this. Here are some possible transformations.
a. Fix missing values (maybe with a Mean or Median value)
b. Create flags to suggest if a variable was missing
c. Transform data by putting it into buckets
d. Mathematical transforms such as log or square root (or use Box-Cox)
e. Combine variables (such as ratios or adding or multiplying) to create new variables



# Build Models
Using the training data set, build at least three different multiple linear regression models, using different variables
(or the same variables with different transformations). Since we have not yet covered automated variable
selection methods, you should select the variables manually (unless you previously learned Forward or Stepwise
selection, etc.). Since you manually selected a variable for inclusion into the model or exclusion into the model,
indicate why this was done.
Discuss the coefficients in the models, do they make sense? For example, if a team hits a lot of Home Runs, it
would be reasonably expected that such a team would win more games. However, if the coefficient is negative
(suggesting that the team would lose more games), then that needs to be discussed. Are you keeping the model
even though it is counter intuitive? Why? The boss needs to know.

\newpage

# Select Models 
Decide on the criteria for selecting the best multiple linear regression model. Will you select a model with slightly
worse performance if it makes more sense or is more parsimonious? Discuss why you selected your model.
For the multiple linear regression model, will you use a metric such as Adjusted R2
, RMSE, etc.? Be sure to
explain how you can make inferences from the model, discuss multi-collinearity issues (if any), and discuss other
relevant model output. Using the training data set, evaluate the multiple linear regression model based on (a)
mean squared error, (b) R2
, (c) F-statistic, and (d) residual plots. Make predictions using the evaluation data set.

# Appendix A

## Data Dictionary

```{r, results='asis', echo=FALSE, cache=TRUE}
kable(data_dictionary)
```

## R code used in document
