---
title: "Homework 1"
author: "Group 1"
date: ""
output: pdf_document
---

base hits, double plays, walks allowed

1. DATA EXPLORATION (25 Points)

```{r, echo=FALSE}
library(moments)
library(RCurl)
x = getURL("https://raw.githubusercontent.com/ChristopheHunt/DATA-621-Group-1/master/imputed-moneyball-training-data.csv")
moneyball = read.csv(text = x, header = T)
newmoneyball = subset(moneyball, select = c(Number.of.wins, Base.Hits.by.batters..1B.2B.3B.HR., Walks.allowed, Double.Plays))
colnames(newmoneyball) = c("Number of Wins", "Base Hits by Batters", "Walks Allowed", "Double Plays")
boxplot(newmoneyball)
```

According to the boxplot shown above, there are lots of outliers in the number of walks allowed, indicating a strong skew to the right. 

```{r, echo=FALSE}
skewness(newmoneyball)
```

The skewness of each variable was obtained. The two variables with a significant amount of skew were the number of base hits by batters and the number of walks allowed. Correspondingly, these two variables had a skew of 1.57 and 6.74. 

```{r, echo=FALSE}
layout(matrix(c(1,2,3,4), 2, 2))
hist(newmoneyball$`Number of Wins`, main = ("Histogram of Number of Wins"), xlab = ("Number of Wins"))
hist(newmoneyball$`Base Hits by Batters`, main = ("Histogram of Base Hits"), xlab = ("Base Hits"))
hist(newmoneyball$`Walks Allowed`, main = ("Histogram of Number of Walks Allowed"), xlab = ("Number of Walks Allowed"))
hist(newmoneyball$`Double Plays`, main = ("Histogram of Double Plays"), xlab = ("Double Plays"))
```

Histograms of all of the variables were plotted so that the distribution of the data could be visualized. In the distribution for the number of walks allowed, only two bars exist due to the excessive number of outliers.

```{r, echo=FALSE}
library(matrixStats)
colSds(as.matrix(newmoneyball))
```

The standard deviations for the base hits and the number of walks allowed are both very high (144.6 and 166.4). This implies that both variables are spread out over a wider range of values.  

2. DATA PREPARATION (25 Points)

```{r, echo = FALSE}
winsorize <- function(x, multiple=2.2)
{
  q <- quantile(x)
  iqr <- IQR(x)
  iqrAdjusted <- iqr*multiple
  
  rangeLow <- q['25%']-iqrAdjusted
  rangeHigh <- q['75%']+iqrAdjusted
  
  x[x<rangeLow] <- min(x[x>rangeLow])
  x[x>rangeHigh] <- max(x[x<rangeHigh])
  
  return(x)  
  
}

newmoneyball$`Number of Wins Winsorized` = winsorize(newmoneyball$`Number of Wins`, multiple = 2.2)
newmoneyball$`Base Hits Winsorized` = winsorize(newmoneyball$`Base Hits by Batters`, multiple = 2.2)
newmoneyball$`Walks Allowed Winsorized` = winsorize(newmoneyball$`Walks Allowed`, multiple = 2.2)
newmoneyball$`Double Plays Winsorized` = winsorize(newmoneyball$`Double Plays`, multiple = 2.2)
boxplot(newmoneyball$`Number of Wins Winsorized`, newmoneyball$`Base Hits Winsorized`, newmoneyball$`Walks Allowed Winsorized`, newmoneyball$`Double Plays Winsorized`)
```

First, the variables were all winsorized. As a result, all of the outliers were changed. 

```{r, echo=FALSE}
library(forecast)
l1 = BoxCox.lambda(newmoneyball$`Number of Wins Winsorized`)
newmoneyball$`Number of Wins Transformed` = BoxCox(newmoneyball$`Number of Wins Winsorized`, l1)
```

```{r, echo=FALSE}
newmoneyball$`Walks Allowed Transformed` = (newmoneyball$`Walks Allowed Winsorized`)^(1/3)
#hist(newmoneyball$`Walks Allowed Transformed`, main = ("Histogram of Cube Root of Walks Allowed"), xlab = ("Cube Root of Walks Allowed"))
newmoneyball$`Double Plays Transformed` = (newmoneyball$`Double Plays Winsorized`)^(1/3)
#skewness(newmoneyball)
```

The variables that were transformed were the number of walks allowed and the number of double plays. The cube roots of these two variables were taken. 

The transformation done on the number of wins was the Box Cox transformation.

As a result of all of these transformations, the adjusted-R-squared was increased to 0.18. 

```{r, echo=FALSE}
fit = lm(`Number of Wins Winsorized` ~ `Base Hits Winsorized` + `Double Plays Winsorized` + `Walks Allowed Winsorized`, data = newmoneyball)
summary(fit)
fit1 = lm(`Number of Wins` ~ `Base Hits by Batters` + `Double Plays` + `Walks Allowed`, data = newmoneyball)
summary(fit1)
fit2 = lm(`Number of Wins Transformed` ~ `Base Hits Winsorized` + `Double Plays Transformed` + `Walks Allowed Transformed`, data = newmoneyball)
summary(fit2)
layout(matrix(c(1,2,3,4), 2, 2))
plot(fit2)
```

```{r, echo=FALSE}
apply(moneyball, 2, function(x) sum(is.na(x))) #to check number of NA's
```

First, two variables that had a really large number of missing data points were eliminated. These variables were the number of times a batter was hit by a pitch and the number of times the runner was caught stealing a base. The other variables had no data points missing or less than 300 data points missing.  

The variables were correlated to the target variable which is the targeted number of wins. The variables were correlated three at a time. In each of those batches, at least one variable was anticipated to have a positive effect and at least one was anticipated to have a negative effect. 

The first coefficients in this model both before and after transformation make sense. As the number of base runs increases, the number of wins also increases. However, the second and third coefficients do not make sense. As the number of double plays increases and as the number of walks allowed decreases, the number of wins is supposed to increase. In the untransformed and transformed models, the number of wins has a negative correlation with the number of double plays and a positive correlation with the number of walks allowed. However, after the transformations were done, the significance of the double plays and the walks allowed increased. 

The one variable that did not need to be transformed was the Base Hits variable. This variable already has a strong correlation with the Target Wins. 

4. SELECT MODELS (25 Points)
Decide on the criteria for selecting the best multiple linear regression model. Will you select a model with slightly
worse performance if it makes more sense or is more parsimonious? Discuss why you selected your model.
For the multiple linear regression model, will you use a metric such as Adjusted R2
, RMSE, etc.? Be sure to
explain how you can make inferences from the model, discuss multi-collinearity issues (if any), and discuss other
relevant model output. Using the training data set, evaluate the multiple linear regression model based on (a)
mean squared error, (b) R2
, (c) F-statistic, and (d) residual plots. Make predictions using the evaluation data set.

