---
title: "Homework 5"
author: "Group 1"
date: ''
output:
  pdf_document:
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
---

```{r programming style guide, echo=FALSE, eval=FALSE}
##Style guide
##USE CAMEL CASING!! https://en.wikipedia.org/wiki/Camel_case, start with lowercase then each word after start with upper case
##upper case for vector, lists, etc. ex. X <- c("yes","no","true","false")
##lower case for scalar/single value, ex. x <- 1
##data frame start with df followed by description, ex. dfEval <- evaluation data frame
##assignments use "<-" not "="
```

```{r library load, include = FALSE, cache=TRUE}
library(pacman)
p_load(Hmisc, xlsx, xtable, knitr, scales, magrittr, tidyverse, stringr, e1071, corrplot, knitcitations, bibtex, missForest, abc, foreach, doParallel, stargazer, forecast, matrixStats, glmulti, leaps, data.table, highlight, car, pracma, boot)
```

\begin{center}
\bigskip
\bigskip
\bigskip
Prepared for:\\
\medskip
Dr. Nathan Bastian\\
\smallskip
City University of New York, School of Professional Studies - Data 621\\
\bigskip
Prepared by:\\
\medskip
Group 1\\ 
\medskip
Senthil Dhanapal\\ 
\smallskip
Yadu Chittampalli\\
\smallskip
Christophe Hunt\\  

\end{center}

\newpage

# Introduction

# Statement of the Problem

# Data Exploration  

## Variables Explained