---
title: Regional Liquor Sales in Des Moines, Iowa
author:
  - name: Christophe Hunt
    email: christophe.hunt@spsmail.cuny.edu
    affiliation: CUNY School of Professional Studies
    footnote: Authors
  - name: Senthil Dhanapal
    email: senthil.dhanapal@spsmail.cuny.edu
    affiliation: CUNY School of Professional Studies
  - name: Yadu Chittampalli
    email: yadu.chittampalli@spsmail.cuny.edu
    affiliation: CUNY School of Professional Studies
address:
  - code: CUNY School of Professional Studies
    address: CUNY School of Professional Studies, Data Analytics, New York, NY
abstract: |
  This is the abstract.

  It consists of two paragraphs
  
  #Keywords:  Liquor Sales, Naive Forecast*

bibliography: mybibfile.bib
output: rticles::elsevier_article
---

```{r programming style guide, echo=FALSE, eval=FALSE}
##Style guide
##USE CAMEL CASING!! https://en.wikipedia.org/wiki/Camel_case, start with lowercase then each word after start with upper case
##upper case for vector, lists, etc. ex. X <- c("yes","no","true","false")
##lower case for scalar/single value, ex. x <- 1
##data frame start with df followed by description, ex. dfEval <- evaluation data frame
##assignments use "<-" not "="
```

```{r options and library load, include = FALSE}
options(xtable.comment = FALSE)
library(pacman)
p_load(ggplot2, fitdistrplus, logspline, pander, Hmisc, tidyverse, stargazer, data.table)
```

Problem
=================

The objective of this report is to create a statistical model for the volume sold of liquor in gallons and the profit dollars in the City of Des Moines which is within the state of Iowa. This can help us make informed decisions on inventory production, sales, and assist wholesale distributors to plan for the predicted volume of distribution.  

Introduction
=================

In February, the Distilled Spirits Council (DISCUS), announced that spirits had an estimated retail sales of nearly $72 billion in 2015. Additionally, DISCUS credits the continuous growth of the distilled spirits industry to several key factors - continuous fascination with American Whiskeys in the United States and abroad, innovations in flavors, permutation across all spirits categories leading to consumer interest, improved regulatory and tax environment resulting in expanded market access and a relatively low number of state tax threats, and the growth of small distillers, which expanded grassroots and overall interest in the spirits category @KeepingSpiritsHigh1. 

This establishes that spirit sales in the Unites States is a valuable market worth exploring for a more detailed and statistical understanding of sales and volume. We hope to more thoroughly understand what impact specific store sights may have accounting for the seasonal impact in November that might effect liquor sales. We will limit the analysis to the City of Des Moines for only whisky sales in the month of November. In 2000 the State of Iowa reported sales at a record pace during the last half of 2000 @IowaSetsRecord2. The later part of the year has an increase in sales so planning to meet capacity is a suitable goal for any company. Our years of interest for this analysis will be the month of November for 2015 and 2016.

Research Background
=================

The main goal that has to be achieved in inventory prediction is increasing the efficiency without decreasing the service value offered to the customers. When managing the levels of inventory, it is important to maintain moderate level(s) - not too high and not too low. If the inventory level is excessive, business funds can get wasted. These funds would not be able to be used for any other purpose, thus involving an opportunity cost. The costs of shortage, handling insurance, recording and inspection would proportionately increase along with inventory volume, thus impairing profitability. 

On the other hand, low level(s) of inventory may result in frequent interruptions in the production schedule resulting in under-utilization of capacity and lower sales. When making predictions about orders that should be placed, assumptions are made as follows - uncertainty always exists regardless of the method(s) used, new technologies cannot always be forecasted for which paradigms do not exist, and social policy will be formulated where the future would be affected, changing the accuracy of the forecast. 

One useful method for predicting inventories is the extrapolation of trends. In this method, trends and cycles in the historical data are examined and mathematical techniques are used to extrapolate to the future. The model chosen for forecasting would depend on the historical data. One of the most common models used in this method is decomposition, where historical data is separated into trend, seasonal, and random components. As a result, forecasts are produced using "turning point analysis". Other examples of models used are adaptive filtering, Box-Jenkins analysis, simple linear regression, curve fitting, and weighted smoothing. 

According to Makridakis, "Judgmental forecasting is superior to mathematical models, although there are several forecasting applications where computer-generated models would be more feasible." When inventory levels for bulk-quantity items would need to be forecasted monthly by large manufacturing companies, generating models through computer software would be more efficient. 

http://www.statpac.org/research-library/forecasting.htm

Forecasting the demand of a product is very essential in predicting the order quantity. As a result, a data bank is created, helping the decision makers settle targets, create plans, and demonstrate changes in the business setting. Two different methods are utilized in the investigation of the future demand - quantitative and qualitative. In quantitative methods, mathematical consistencies in the history are searched for. Two subcategories exist in quantitative methods - time series models and correlation models. On the other hand, qualitative methods are based on the opinions that people have had about the product in the past based on their experiences, premonitions, and emotions. 

However, when the most suitable forecast model gets selected, it is not necessarily based solely on quantitative or qualitative variables. The forecast model can even combine several models. 

http://www.slideshare.net/AbhishekKumar159/demand-forecast-process-and-inventory-management


There was a study conducted by the Beer Industry Electronic Commerce Coalition. The purpose of this study was to highlight, explain, and create future considerations for the advancement of forecasting future beer sales and the process around how distributors and suppliers can collaborate to achieve this goal. Weekly forecasts can be created by supplier, brand, package, etc. for up to fifteen weeks into the future if necessary. It is important to be aware of the product trends so that specific situations can be explained to various employees of the company at any time.

https://www.nbwa.org/sites/default/files/hitting_the_mark.pdf

Methodology
================

Our initial data set is sufficiently large in that it includes sales by individual stores and the invoices for each store. The reason for the large size of the initial data set is due to it including every liquor transaction from 2012 to present in Iowa, so it approaches 2.68 GB. For the purposes of this analysis, to analyze a data set this large is not feasible. Therefore, we reduced the number of variables and summarized to a regional aggregate.

Additionally, we looked into the top 10 liquor categories for each year by number of bottles sold. In 2015, the top categories were American Cocktails, Blended Whiskies, Canadian Whiskies, Imported Vodka, Puerto Rico & Virgin Islands Rum, Spiced Rum, Straight Bourbon Whiskies, Tequila, Vodka 80 Proof, and Whiskey Liqueur. Interestingly straight bourbon appears to have more sales in 2015 than 2014 which coincides with the literature of strong growing whiskey sales for every whiskey segment (@SpecialityGrow3). We decided to focus on whiskey do to its strong sales and growing interest in the US.

Furthermore, we looked into the sales by `County` for Iowa 

We initially attempted to model for a dependent variable of `Volume Sold in Gallons` by the top Counties and top liquor categories, however, the distribution of this variable becomes over-dispersed and negatively skewed when aggregating the data set and therefore we were unsuccessful in modeling this variable across regions. Our hope was that if we could model for the gallons sold by entirety of Iowa we could more accurately predict our planned inventory and anticipate production goals. We were more successful in limiting our regional analysis to one city and the stores within that city. 

Therefore, our final evaluation data set is the following subset of variables for largest city in Iowa of Des Moines as follows; `Vendor Name`, `Pack` (pack size of bottles sold) `Bottle Volume`, `State Bottle Cost`, `State Bottle Retail`, `Sales Dollars` (Total sales), and our dependent variables are `Volume sold in Gallons` and `Profit Dollars`.


We then sought to model the mean price per bottle, by modeling the price per bottle we can make pricing decisions based on region and the volume we plan to sell. 


discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection.

Experimentation and Results
=============================

### Data Acquisition

The data set contains the spirits purchase information of Iowa Class "E" liquor licensees by product and date of purchase from January 1, 2012 to current. The data set is provided by the Iowa Department of Commerce, Alcoholic Beverages Division, [click here](https://data.iowa.gov/Economy/Iowa-Liquor-Sales/m3tr-qhgy) to view the data set at Data.Iowa.Gov.

As previously discussed, the data set is 2.68 GB in total size and much to large to use in a meaningful model.

```{r, echo = FALSE, results='asis', warning = FALSE, message = FALSE, cache=TRUE}
iowaDataReduced <- read.csv("https://raw.githubusercontent.com/ChristopheHunt/DATA-621-Group-1/master/Final%20Project/iowa_data_reduced.csv") 
```

```{r, echo = FALSE, results='asis', warning = FALSE, message = FALSE, cache=TRUE}
library(magrittr)
library(dplyr)

iowaDataReducedTop <- iowaDataReduced %>%
                      dplyr::group_by(Year, Category.Name) %>%
                      dplyr::select(Year, Category.Name, Volume.Sold..Gallons. ) %>%
                      dplyr::summarize(Volume.Sold..Gallons. = round(sum(Volume.Sold..Gallons.))) %>%
                      dplyr::top_n(n = 10, wt = Volume.Sold..Gallons.)

top10in2015 <- iowaDataReducedTop %>%
               dplyr::filter(Year == 2015)

top20counties <- iowaDataReduced %>%
                  dplyr::filter(Year == 2015) %>% 
                      dplyr::group_by(Year, County) %>%
                      dplyr::select(Year, Category.Name, Volume.Sold..Gallons. ) %>%
                      dplyr::summarize(Volume.Sold..Gallons. = sum(Volume.Sold..Gallons.)) %>%
                      top_n(n = 20, wt = Volume.Sold..Gallons.)
```

```{r, echo = FALSE, results='asis', warning = FALSE, message = FALSE, cache=TRUE}
library(magrittr)

iowaDataReduced <- iowaDataReduced %>%
                   filter(Year == 2015) %>%
                   dplyr::select(County, Category.Name, 
                                 Bottles.Sold, Sale..Dollars., Volume.Sold..Gallons.) %>%
                  mutate(`Price per Bottle` = Sale..Dollars./Bottles.Sold) %>%
                   dplyr::group_by(County, Category.Name) %>%
                   dplyr::summarize(`Bottles Sold` = sum(Bottles.Sold), 
                             `Avg Price Per Bottle` = mean(`Price per Bottle`),
                             `Sales in Dollars` = sum(Sale..Dollars.), 
                             `Volume Sold by Gallons` = sum(Volume.Sold..Gallons.)) %>%
                  filter(County != "") %>%
                  filter(Category.Name %in% top10in2015$Category.Name) %>%
                  filter(County %in% top20counties$County)
```

```{r, echo = FALSE, eval=FALSE}
write.csv(iowaDataReduced, "iowa_data_reduced.csv")
```

We reviewed the liquor sales by gallons sold per year by Liquor Category. Initially, we viewed the top 5 Liquor Categories by volume sold but there were large disparaties between years, suggesting that the top 5 change often and is likely due to changing consumer tastes. We do see a more stable set of liquor categories for the top 10 category which suggests that while tastes may change we don't see large movements in liquor categories at this level. We focused our attention on the whiskey categories.  

```{r, echo = FALSE, results='asis', warning = FALSE, message = FALSE, cache=TRUE, fig.height=5, fig.width=10, eval=TRUE}
ggplot(iowaDataReducedTop, aes(x = Category.Name, y = Volume.Sold..Gallons., fill = factor(Year))) + 
      scale_y_continuous(labels = scales::comma) + 
      geom_bar(stat = "identity") +
      facet_grid(~Year) +
      theme_minimal() + 
      theme(axis.text.x = element_text(angle = 70, hjust = 1), legend.position = "none") + 
      coord_flip() + 
      xlab("Liquor Category") + 
      ylab("Volume Sold by Gallons\n") +
      ggtitle("Volume of Liquor Category Sales by Gallons Sold")
```

We can further see that Des Moines accounts for a significant portion of the liquor sales in Polk County. Polk County is the most populous county in Iowa so we will limit our analysis to this city.

```{r, echo = FALSE, results='asis', warning = FALSE, message = FALSE, cache=TRUE, eval=FALSE}

iowaCounties   <- iowa_data %>%
                     dplyr::select(Date, City, County, `Category Name`, `Store Number`, `Store Name`, `Pack`, 
                            `Item Number`, `Item Description`, `Bottle Volume (ml)`, 
                            `State Bottle Cost`, `State Bottle Retail`, `Sale (Dollars)`,
                             `Volume Sold (Gallons)`, `Bottles Sold`) %>%
                    mutate(Date = as.Date(Date, format = "%m/%d/%Y"),
                           City = factor(toupper(City)),
                           County = factor(toupper(County)),
                           `Category Name` = factor( toupper(`Category Name`)),
                           `Store Number` = factor(`Store Number`),
                           `Store Name` = factor(`Store Name`), 
                           `Pack` = as.integer(`Pack`),
                           `Bottle Volume (ml)` = as.numeric(`Bottle Volume (ml)`), 
                           `State Bottle Cost` = removeDollars(`State Bottle Cost`), 
                           `State Bottle Retail` = removeDollars(`State Bottle Retail`), 
                           `Sale (Dollars)` = removeDollars(`Sale (Dollars)`),
                           `Volume Sold (Gallons)` = as.numeric(`Volume Sold (Gallons)`), 
                           `Bottles Sold` = as.integer(`Bottles Sold`)) %>%
                    mutate(Month = format(Date, "%m"),
                           Year = format(Date, "%Y")) %>%
                    dplyr::filter(grepl("WHISKIES", `Category Name`)) %>%
                    dplyr::select(-Date) %>%
                    group_by(Month, Year, City, County, `Category Name`, `Store Number`, `Store Name`)  %>%
                    summarize( `Pack` = mean(`Pack`),
                               `Bottles Sold` = sum(`Bottles Sold`), 
                               `Sale (Dollars)` = sum(`Sale (Dollars)`), 
                               `Bottle Volume (ml)` = mean(`Bottle Volume (ml)`),
                               `State Bottle Cost` = sum(`State Bottle Cost`),
                               `State Bottle Retail` = sum(`State Bottle Retail`),
                               `Sale (Dollars)` = sum(`Sale (Dollars)`),
                               `Volume Sold (Gallons)` = sum(`Volume Sold (Gallons)`)) %>%
                    as_tibble()

write.csv(iowaCounties, "iowaCounties.csv")
```

```{r echo = FALSE, results='asis', warning = FALSE, message = FALSE, cache=TRUE, eval=TRUE}
iowaCounties <- fread("https://raw.githubusercontent.com/ChristopheHunt/DATA-621-Group-1/master/Final%20Project/Final%20Project%20Markdown%20Folder/iowaCounties.csv", showProgress = FALSE)

iowaCounties <- iowaCounties %>% 
                ungroup() %>%
                mutate(County = factor(County),
                       City = factor(City)) %>% 
                dplyr::filter(!(County == '')) %>%
                select(Year, County, City, `Volume Sold (Gallons)`) %>%
                group_by(Year, County, City) %>%
                summarize(`Volume Sold (Gallons)` = sum(`Volume Sold (Gallons)`))
```

```{r, echo = FALSE, results='asis', warning = FALSE, message = FALSE, cache=TRUE, fig.height=4.05, fig.width=10, eval=TRUE}
ggplot(iowaCounties %>% filter(County == "POLK"), aes(x = County, y = `Volume Sold (Gallons)`, fill = City)) + 
      scale_y_continuous(labels = scales::comma) + 
      geom_bar(stat = "identity") +
      facet_grid(~Year) +
      theme_minimal() + 
      theme(axis.text.x = element_text(angle = 70, hjust = 1)) + 
      guides(fill = guide_legend(keywidth = 2, keyheight = .75)) + 
      xlab("County") + 
      ylab("Volume Sold by Gallons\n") +
      ggtitle("Gallons Sold by Polk County and City for Whiskey Categories")
```

Our first attempt was to  use a Poisson regression due to the over-dispersion created by aggregating the data set. However, the distribution was far to negatively skewed to fit a poission distribution. 

## Distributions of Dependent Variables

```{r, echo = FALSE, results='asis', warning = FALSE, message = FALSE, cache=TRUE, fig.height=6, fig.width=10, eval=TRUE}
x <- descdist(iowaDataReduced$`Bottles Sold`, discrete = FALSE, boot = 100)
pandoc.table(x)
```

```{r, echo = FALSE, results='asis', warning = FALSE, message = FALSE, cache=TRUE, fig.height=6, fig.width=10, eval=TRUE}
fit.norm <- fitdist(iowaDataReduced$`Bottles Sold`, "norm")
plot(fit.norm)
```


describe the specifics of what you did (data exploration, data preparation,
model building, model selection, model evaluation, etc.), and what you found out (statistical analyses,
interpretation and discussion of the results, etc.).


```{r , echo = FALSE, results='asis', warning = FALSE, message = FALSE , cache=TRUE, eval=FALSE}
library(data.table)
library(tidyverse)

iowa_data <- fread("~./Iowa_Liquor_Sales.csv", header = T, sep = ',')

removeDollars <- function(column){
  as.numeric(gsub("\\,","",gsub("\\$","", column)))
}

iowa_data_reduced <- iowa_data %>%
                     dplyr::select(Date, City, `Category Name`, `Store Number`, `Store Name`, `Pack`, 
                            `Item Number`, `Item Description`, `Bottle Volume (ml)`, 
                            `State Bottle Cost`, `State Bottle Retail`, `Sale (Dollars)`,
                             `Volume Sold (Gallons)`, `Bottles Sold`) %>%
                    mutate(Date = as.Date(Date, format = "%m/%d/%Y"),
                           City = factor(toupper(City)),
                           `Category Name` = factor( toupper(`Category Name`)),
                           `Store Number` = factor(`Store Number`),
                           `Store Name` = factor(`Store Name`), 
                           `Pack` = as.integer(`Pack`),
                           `Bottle Volume (ml)` = as.numeric(`Bottle Volume (ml)`), 
                           `State Bottle Cost` = removeDollars(`State Bottle Cost`), 
                           `State Bottle Retail` = removeDollars(`State Bottle Retail`), 
                           `Sale (Dollars)` = removeDollars(`Sale (Dollars)`),
                           `Volume Sold (Gallons)` = as.numeric(`Volume Sold (Gallons)`), 
                           `Bottles Sold` = as.integer(`Bottles Sold`)) %>%
                    mutate(Month = format(Date, "%m"),
                           Year = format(Date, "%Y")) %>%
                    dplyr::filter(Month == "11") %>%
                    dplyr::filter(Year == "2015") %>% 
                    dplyr::filter(City == "DES MOINES") %>% 
                    dplyr::filter(grepl("WHISKIES", `Category Name`)) %>%
                    dplyr::select(-Date) %>%
                    group_by(Month, Year, City, `Category Name`, `Store Number`, `Store Name`)  %>%
                    summarize( `Pack` = mean(`Pack`),
                               `Bottles Sold` = sum(`Bottles Sold`), 
                               `Sale (Dollars)` = sum(`Sale (Dollars)`), 
                               `Bottle Volume (ml)` = mean(`Bottle Volume (ml)`),
                               `State Bottle Cost` = sum(`State Bottle Cost`),
                               `State Bottle Retail` = sum(`State Bottle Retail`),
                               `Sale (Dollars)` = sum(`Sale (Dollars)`),
                               `Volume Sold (Gallons)` = sum(`Volume Sold (Gallons)`)) %>%
                    as_tibble()

iowa_data_reduced <- iowa_data_reduced %>%
                    dplyr::ungroup() %>%
                    mutate(Month = as.numeric(Month),
                           Year = as.numeric(Year),
                           `Category Name`= factor(`Category Name`)) %>%
                    arrange(Year, Month, County, City)

write.csv(iowa_data_reduced, "iowa_data_reduced.csv")
```


# Model 1 

```{r data read for model, echo = FALSE, cache=TRUE}
dfLiquorSales <- read.csv("https://raw.githubusercontent.com/ChristopheHunt/DATA-621-Group-1/master/Final%20Project/iowa_data_reduced%2012.10.2016.csv") %>% mutate(Store.Number = factor(Store.Number))
```


#Forward step 

```{r forward step selection for model, echo = FALSE, cache=TRUE, include=FALSE} 
step( lm(Bottles.Sold~1, data = dfLiquorSales), direction = "forward", scope=~Category.Name + Store.Number + Pack + Sale..Dollars. + Bottle.Volume..ml. + State.Bottle.Cost + State.Bottle.Retail + Volume.Sold..Gallons., trace = FALSE)
```

```{r model 1, echo = FALSE, cache=TRUE, results = 'asis'}
fit1 <- lm(formula = Bottles.Sold ~ Volume.Sold..Gallons. + Category.Name +  State.Bottle.Retail + Pack + Sale..Dollars., data = dfLiquorSales)
stargazer(fit1, header = FALSE, no.space = TRUE, style = "all2", font.size = "normalsize", single.row = TRUE, intercept.bottom = FALSE, title = "Forward Selection Linear Model 1")
```


#Volume.Sold..Gallons=One unit of increase in volume sold will increase the bottles sold by 4.10 units

#Category.Name
# CANADIAN WHISKIES is the only one that is significant. 
# CANDIAN WHISKIES vs BLENDED WHISKIES coefficient is 5.492

#State.Bottle.Retail=One unit of increase in State.Bottle.Retail will increase the bottles sold by 7.133 units
#Pack=One unit of increase in Pack will increase the bottles sold by 7.133 units
#Sale..Dollars.=It is slightly significant. One unit of increase in Sale..Dollars. will decrease the bottles sold by 2.189 units

#Formula = -4.196 + 25.314*Volume.Sold..Gallons. + 5.492 * (Category.Name="CANADIAN WHISKIES" ) + 
#       7.133 * (State.Bottle.Retail) + 7.521 * (Pack) - 2.189 * Sale..Dollars. 




Discussion and Conclusions
============================

In two studies that were conducted, the statistical data from the previous time interval was used to predict what would happen in the next time interval. One study was for pharmaceutical distribution companies. The other study involved data mining models.  

In one study involving pharmaceutical distribution companies, the purpose was to propose a novel method to forecast the sales of the companies. Network-based analysis was conducted to find clique sets and group members and to use the sales data of comembers. The reason for this was the lack of sufficient historic sales records of each drug. 

Three methods were used to build time series models forecasting sales - ARIMA methodology, neural network, and an advanced hybrid neural network approach. The performance of the proposed method was evaluated using a real dataset provided by one of the leading pharmacy distribution companies in Iran. The results of the evaluation indicated that the proposed method can cope with the low number of past records while accurately forecasting medicine sales.

After exploratory analysis was done on the data, it was concluded that most medicines had different and specific characteristics and sales behavior, it was impractical to make a single prediction model for all medicines, and most sales records had nonlinear relationships. 

The reason why the hybrid neural network method was carried out was due to the fact that it is not acceptable to apply a fully linear or nonlinear model on sales data. 

The two forecast error measures that were used to evaluate and compare model performance were mean squared error and mean absoute error. The performance of the predicted data was significantly improved when the past records of comembers were used. 

http://onesearch.cuny.edu/primo_library/libweb/action/display.do?tabs=detailsTab&ct=display&fn=search&doc=TN_gale_ofa425455991&indx=3&recIds=TN_gale_ofa425455991&recIdxs=2&elementId=2&renderMode=poppedOut&displayMode=full&frbrVersion=2&frbg=&&dscnt=0&scp.scps=scope%3A%28CUNY_BEPRESS%29%2Cscope%3A%28BB%29%2Cscope%3A%28BB_LIBGUIDES%29%2Cscope%3A%28AL%29%2Cprimo_central_multiple_fe&tb=t&mode=Basic&vid=bb&srt=rank&tab=default_tab&dum=true&vl(freeText0)=inventory%20prediction%20sales&dstmp=1481346949259

The other study that examined prediction-based inventory optimization using data mining models, the Back propagation neural network was used for training the prediction model. The idea that gave rise to this method of inventory prediction was the idea that the demand of marketing is viewed as the foundation of inventory management.   

On the basis of the prediction result, a simple and concise inventory policy was established. Following this, the historic sales data was used to estimate a normal distribution of demand and to calculate the inventory cost with inventory strategy. 

Two models (Back Propagation Neural Network and Support Vector Regression) were established using three input variables (historical sales data, the frequency of searching the commodity, and the click volume of the commodity page).

When the back-propagation neural network method was used, there was more accuracy shown in the performance because the predicted values almost matched the actual values in the graphs.

The mean absolute percentage error calculated for this model is 0.06. 

http://onesearch.cuny.edu/primo_library/libweb/action/display.do?frbrVersion=2&tabs=detailsTab&ct=display&fn=search&doc=TN_ieee10.1109%2fCSO.2014.118&indx=10&recIds=TN_ieee10.1109%2fCSO.2014.118&recIdxs=9&elementId=9&renderMode=poppedOut&displayMode=full&frbrVersion=2&frbg=&&dscnt=0&scp.scps=scope%3A%28CUNY_BEPRESS%29%2Cscope%3A%28BB%29%2Cscope%3A%28BB_LIBGUIDES%29%2Cscope%3A%28AL%29%2Cprimo_central_multiple_fe&tb=t&mode=Basic&vid=bb&srt=rank&tab=default_tab&dum=true&vl(freeText0)=inventory%20prediction%20sales&dstmp=1481412549638


In another study conducted in 2012 in Idaho, the monthly revenue generated was examined rather than the yearly revenue generated. The continued growth was rather owed to the number of weekends a month has (five instead of four) and to the higher prices in neighboring states. In Washington, the voters approved an initiative that led the state to sell its liquor stores and add new distributor and retail fees, making prices in the neighboring states (Idaho and Oregon) look better. There were no changes made in marketing or pricing in response to the regulatory shift in Washington @IdahoSales4. Further research into the proximity of our counties to states and towns with higher prices and regulation may provide more insight into sales and volume of liquor sold. Additionally, reviewing the data by identifying months that has 5 weekends instead of four could provide further insights. 

conclude your findings, limitations, and suggest areas for future work.

\newpage

Appendices
=============

# Supplemental tables and/or figures.

```{r data set decsribed, results='asis', echo=FALSE}
describedData <- describe(dfLiquorSales)
latex(describedData, file = '')
```

```{r}
describe(dfLiquorSales$Store.Number)
describe(dfLiquorSales$Category.Name)
```

# Session Info

```{r, results='asis', echo=FALSE, eval = TRUE}
toLatex(sessionInfo())
```

# R statistical programming code.

Please see [Final Project.rmd](https://github.com/ChristopheHunt/DATA-621-Group-1/blob/master/Final%20Project/Final%20Project.Rmd) on GitHub for source code.   

https://github.com/ChristopheHunt/DATA-621-Group-1/blob/master/Final%20Project/Final%20Project.Rmd

References {#references .unnumbered}
===============
