---
title: "Homework 2"
author: "Group 1"
date: ''
output:
  pdf_document:
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
---

\begin{center}
\bigskip
\bigskip
\bigskip
Prepared for:\\
\medskip
Dr. Nathan Bastian\\
\smallskip
City University of New York, School of Professional Studies - Data 621\\
\bigskip
Prepared by:\\
\medskip
Group 1\\ 
\medskip
Senthil Dhanapal\\ 
\smallskip
Yadu Chittampalli\\
\smallskip
Christophe Hunt\\  

\end{center}

```{r, include = FALSE}
library(pacman)
p_load(xtable, knitr, pander, xtable, scales, tidyverse, formatR)
```


# Data Source 

We download the data from our public GitHub repository which was originally provided through Blackboard. 

```{r, echo=FALSE}
classification <- read.csv("https://raw.githubusercontent.com/ChristopheHunt/DATA-621-Group-1/master/Homework%202/classification-output-data.csv")
```

# Data Explained

We will be using the following columns from the data source: 

* class: the actual class for the observation
* scored.class: the predicted class for the observation (based on a threshold of 0.5)
* scored.probability: the predicted probability of success for the observation

Below is the raw confusion matrix for our scored dataset

```{r, echo=FALSE, results = 'asis', cache=TRUE}
scores <- subset(classification, select = c("class", "scored.class", "scored.probability"))
m <- table("Predicted" = scores$class, "Actual" = scores$scored.class)
colnames(m) <- c('Predicted Failure', "Predicted Success")
rownames(m) <- c('Actual Failure', 'Actual Success')
kable(m)
```

In particular, do the rows represent the actual or predicted class? The columns? - Yadu

# Accuracy of Predictions 

We developed the below function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the accuracy of the predictions. - Yadu 

```{r, results='asis'}
accuracy <- function(df, actual, prediction){
  require(scales)
  truefalse <- as.data.frame(table(df[actual] == df[prediction])) %>% 
               spread('Var1', 'Freq')
  accuracy <- unlist(truefalse[2]/nrow(classification))
  paste0("The prediction accuracy is ", percent(accuracy))
}

accuracy(df = scores, actual = 'class', prediction = 'scored.class')
```

4. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the classification error rate of the predictions.Verify that you get an accuracy and an error rate that sums to one. - Yadu 

5. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the precision of the predictions. - Senthil

6. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the sensitivity of the predictions. Sensitivity is also known as recall. - Senthil

7. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the specificity of the predictions. - Christophe

8. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the F1 score of the predictions. - Christophe

9. Before we move on, let's consider a question that was asked: What are the bounds on the F1 score? Show
that the F1 score will always be between 0 and 1. (Hint: If 0 < ???? < 1 and 0 < ???? < 1 then ????????- Christophe

10. Write a function that generates an ROC curve from a data set with a true classification column (class in our
example) and a probability column (scored.probability in our example). Your function should return a list
that includes the plot of the ROC curve and a vector that contains the calculated area under the curve
(AUC). Note that I recommend using a sequence of thresholds ranging from 0 to 1 at 0.01 intervals. - Senthil

11. Use your created R functions and the provided classification output data set to produce all of the
classification metrics discussed above. - Christophe

12. Investigate the caret package. In particular, consider the functions confusionMatrix, sensitivity, and
specificity. Apply the functions to the data set. How do the results compare with your own functions? - Yadu

13. Investigate the pROC package. Use it to generate an ROC curve for the data set. How do the results
compare with your own functions? - Senthil 
