---
title: "Homework 4"
author: "Group 1"
date: ''
output:
  pdf_document:
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
---

```{r programming style guide, echo=FALSE, eval=FALSE}
##Style guide
##USE CAMEL CASING!! https://en.wikipedia.org/wiki/Camel_case, start with lowercase then each word after start with upper case
##upper case for vector, lists, etc. ex. X <- c("yes","no","true","false")
##lower case for scalar/single value, ex. x <- 1
##data frame start with df followed by description, ex. dfEval <- evaluation data frame
##assignments use "<-" not "="
```

```{r library load, include = FALSE}
library(pacman)
p_load(Hmisc, xlsx, xtable, knitr, scales, magrittr, tidyverse, stringr, e1071, corrplot, knitcitations, bibtex, missForest, abc, foreach, doParallel, stargazer, forecast, matrixStats, glmulti, leaps, data.table, highlight, car)
```

\begin{center}
\bigskip
\bigskip
\bigskip
Prepared for:\\
\medskip
Dr. Nathan Bastian\\
\smallskip
City University of New York, School of Professional Studies - Data 621\\
\bigskip
Prepared by:\\
\medskip
Group 1\\ 
\medskip
Senthil Dhanapal\\ 
\smallskip
Yadu Chittampalli\\
\smallskip
Christophe Hunt\\  

\end{center}

# Introduction

# Statement of the Problem

# Data Exploration  

## Variables Explained

The variables provided in our evaluation data set our explained below:

```{r data dictionary table, eval=FALSE, results='asis', echo=FALSE}
kable(daDict %>% 
      filter(VARIABLE.TYPE == "Predictor Variable") %>%
      mutate(Abbreviation = VARIABLE.NAME, Definition = DEFINITION)  %>%
      dplyr::select(Abbreviation, Definition),
      align = c("c","l"))
```

Histograms of most of our variables have been plotted below so that distribution can be visualized. 

```{r descriptive statistics table, echo=FALSE, results='asis', cache=TRUE}

dfTr <- read.csv("https://raw.githubusercontent.com/ChristopheHunt/DATA-621-Group-1/master/Homework%204/insurance_training_data.csv")

digits <- c(1)

descriptive <- describe( dfTr %>% dplyr::select(-INDEX, -TARGET_FLAG) 
                        , descript = "Table 1 : Descriptive Statistics", digits = digits)
latex(descriptive, file = '')
```

\newpage

## Exploration of Variables

```{r correlation matrix, fig.cap= "Correlation Plot of Training Data Set", echo = FALSE, fig.height = 3, fig.width= 8, cache=TRUE}

removedollars <- function(column){
             as.numeric(gsub("\\,","",gsub("\\$","", column)))
          }

dfTr <- dfTr %>%
        mutate(INCOME = removedollars(INCOME),
               HOME_VAL = removedollars(HOME_VAL),
               BLUEBOOK = removedollars(BLUEBOOK),
               OLDCLAIM = removedollars(OLDCLAIM),
               MSTATUS = gsub("z_", "", MSTATUS),
               SEX = gsub("z_", "", SEX))



dfTrMx <- as.matrix(dfTr)
corMx <- cor(dfTrMx , use = "everything",  method = c("pearson"))
corrplot(corMx, order = "hclust", addrect = 2, method = "square", tl.col = "black", tl.cex = .75, na.label = " ")
```


